\textbf{Supplementary materials: Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}

### 2024-01-03 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Material and methods--- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and methods}

@@

%%% ---------------------------------------------------------------- %%%
%%% Table 1. Outdoor air pollution concentration BiSC-home, BiSCAPE  %%%
%%% ---------------------------------------------------------------  %%%
\begin{table}[h!]
\centering
\caption{Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns (2018-2022)}
\label{table1} 
\begin{tabular}{lllllllllll} % each one indicates a new column in the table 
\toprule
Campaign & Pollutant & Year  & mean & sd & min & p25 & p50 & p75 & max \\ 
\midrule
BiSC-home, BiSCAPE & NO\textsubscript{2} (\mu\)g/m\textsuperscript{3}) & 2018-2021 & 38.8 & 12.4 & 6.85 &30.0 & 37.5 & 45.8 & 97.3 \\
\hline
\multirow{5}{*}{BiSCAPE} 
& PM\textsubscript{2.5} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 13.4 & 6.14 & 5.01 & 9.06 & 11.3 & 16.2 & 35.8 \\ 
& BC (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 1.27 & 0.50 & 0.40 & 0.88 & 1.22 & 1.52 & 2 .95 \\ 
& PM\textsubscript{2.5 - Fe} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 0.25 & 0.13 & 0.07 & 0.15 & 0.21 & 0.32 & 0.77 \\ 
& PM\textsubscript{2.5 - Cu} (ng/m\textsuperscript{3}) & 2021-2022 & 7.10 & 3.82 & 1.78 & 4.37 & 6.36 & 8.98 & 22.7 \\ 
& PM\textsubscript{2.5 - Zn} (ng/m\textsuperscript{3}) & 2021-2022  & 40.7 & 36.5 & 6.71 &  18.9 & 28.2 & 46.7 & 273 \\ 
\bottomrule
\end{tabular}
\end{table}

@@

%%% -------------------------------------------%%%%
%%% --- Figure 2. Boxplot models comparison --- %%%
%%% -------------------------------------------%%%%
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/boxplot_all_models_estimates.png}
\caption{Distribution of exposure to outdoor air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
\label{fig2}
\end{figure}

@@

%%% --- Figure. Average exposure during the entire pregnancy period --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/all_maps.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
\end{figure}

@@

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Discussion --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
We successfully developed spatiotemporal models using three different approaches (LUR, DM, and HM). Models show moderate to good performance

spatiotemporal models using three different approaches (LUR, DM and HM). All models show moderate to good performance evaluated by the defined metrics in the \textbf{methods section} and displayed in more detail in the \textbf{Supplementary materials}.

As we expected, hybrid models outperformed the LUR and DM models for the air pollutants assessed. We believe that by combining multiple data sources, and integrating RF algorithm, we were able to capture non-linear and interactions between variables and consequently improve model predictive ability.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Material and methods--- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and methods}

@@

%%% ---------------------------------------------------------------- %%%
%%% Table 1. Outdoor air pollution concentration BiSC-home, BiSCAPE  %%%
%%% ---------------------------------------------------------------  %%%
\begin{table}[h!]
\centering
\caption{Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns (2018-2022)}
\label{table1} 
\begin{tabular}{lllllllllll} % each one indicates a new column in the table 
\toprule
Campaign & Pollutant & Year  & mean & sd & min & p25 & p50 & p75 & max \\ 
\midrule
BiSC-home + BiSCAPE & NO\textsubscript{2} (\mu\)g/m\textsuperscript{3}) & 2018-2021 & 38.8 & 12.4 & 6.85 &30.0 & 37.5 & 45.8 & 97.3 \\
\hline
\multirow{5}{*}{BiSCAPE} 
& PM\textsubscript{2.5} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 13.4 & 6.14 & 5.01 & 9.06 & 11.3 & 16.2 & 35.8 \\ 
& BC (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 1.27 & 0.50 & 0.40 & 0.88 & 1.22 & 1.52 & 2 .95 \\ 
& PM\textsubscript{2.5 - Fe} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 0.25 & 0.13 & 0.07 & 0.15 & 0.21 & 0.32 & 0.77 \\ 
& PM\textsubscript{2.5 - Cu} (ng/m\textsuperscript{3}) & 2021-2022 & 7.10 & 3.82 & 1.78 & 4.37 & 6.36 & 8.98 & 22.7 \\ 
& PM\textsubscript{2.5 - Zn} (ng/m\textsuperscript{3}) & 2021-2022  & 40.7 & 36.5 & 6.71 &  18.9 & 28.2 & 46.7 & 273 \\ 
\bottomrule
\end{tabular}
\end{table}

@@

%%% -------------------------------------------%%%%
%%% --- Figure 2. Boxplot models comparison --- %%%
%%% -------------------------------------------%%%%
\begin{figure}[!htb]
\includegraphics[width=0.9\textwidth]{figures/boxplot_all_models_estimates.png}
\caption{Distribution of exposure to outdoor air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
\label{fig2}
\end{figure}

@@

%%% --- Figure. Average exposure during the entire pregnancy period --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=0.9\textwidth]{figures/combined_figure1.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
\end{figure}

\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=0.9\textwidth]{figures/combined_figure2.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and hm(\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Discussion --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
We successfully developed spatiotemporal models using three different approaches (LUR, DM, and HM). Models show moderate to good performance 
evaluated by the defined metrics previously and displayed in more detail in the \textbf{supplementary materials}. As we expected, hybrid models outperformed the LUR and DM models for the air pollutants assessed. We believe that by combining multiple data sources, and integrating RF algorithm, we were able to capture non-linear and interactions between variables and consequently improve model predictive ability.

### 2024-01-03 supplementary_BiSC_outdoor_AP.tex
Updates from Overleaf

\begin{enumerate}
    \item Define the baseline period, that is, the total period in which the air pollution measurements were done. For NO\textsubscript{2} the baseline period was from 23\textsuperscript{rd} October 2018 to 16\textsuperscript{th} February 2022.
    \item Collect hourly air pollution data from a routine monitoring site covering the baseline period. We used data from Palau Reial monitoring station, as this is the reference site for the BiSCAPE campaigns and the BiSC cohort. NO\textsubscript{2}, data at Palau Reial was missing for only 5 \% of the period. Even though the missing rate is low, we imputed missing data by doing a linear regression with NO\textsubscript{2} levels at the monitoring site of Sants (r = 0.84). After recovering those data, the percentage of missing data at Palau Reial decreased to 0.68 \%. 
    \item Calculate the average concentration of NO\textsubscript{2} for the routine monitoring site covering the baseline measurement period: C\textsubscript{PR}(avg).
    \item Calculate the concentrations of the pollutant at the routine monitoring site for the exact same period (t) as the sample collected at the BiSC and BiSCAPE locations in the different periods: CP\textsubscript{PR}(t). 
    \item Calculate for the routine monitoring site for each sample the difference between the concentration (C\textsubscript{PR}) and the average covering the baseline measurement period: DC\textsubscript{PR}(t) = C\textsubscript{PR}(t) - C\textsubscript{PR}(avg).
    \item Calculate for each sample at site i (C\textsubscript{i}(t)) the adjusted concentration (C\textsubscript{i, adj}(t)) by computing: C\textsubscript{i,adj}(t) = C\textsubscript{i}/DC\textsubscript{PR}(t).
\end{enumerate}

\subsubsection{PM\textsubscript{2.5}, PM\textsubscript{2.5} components (Fe, Cu, Zn), and BC deseasonalization}

For the PM\textsubscript{2.5} data (BiSCAPE + BiSC) the \textit{difference method} was used for deseasonalisation.

@@

@@

@@

@@

\begin{table}[ht]

@@

@@

@@

 & & & & & avg\_traffic\_stations + avg\_solar\_radiation +\\ 

@@

\multirow{4}{*}{PM\textsubscript{2.5}} & \multirow{4}{*}{2021} & \multirow{4}{*}{161} & \multirow{4}{*}{0.66} & \multirow{4}{*}{3.45} & pm25\_dispersion + idw\_pm25monitoring\_stations + \\
& & & & & avg\_traffic\_stations + ldres1000 + build25 +  pop1000 + \\
& & & & & roads25 + sqralt +  majorroadlength25 + \\
& & & & & trafload300 + build\_height\_25 + lat + lon \\

@@

BC & 2021 & 74 & 0.86 &  0.23 & bc\_Dispersion + idw\_nox\_monitoring\_stations +
 avg\_atmospheric\_pressure + avg\_bc\_palau\_reial +  pop100 + hdres300 + roads25 + avg\_wind\_speed + roadlength25 +  hdres50 + distinvmajor1 + avg\_traffic\_stations +  linesnear +  pop300 + 
 trafload500 + roads500
 \\

@@

PM\textsubscript{2.5 - Fe} & 2021 & 161 & 0.54 & 0.08 & pm25\_road\_nonexhaust +idw\_pm25monitoring\_stations + roads300 + roads1000 + ldres500 + sqralt + zbe + lat + lon \\

@@

PM\textsubscript{2.5 - Cu} & 2021 & 154 & 0.70 & 2.15 & pm25\_road\_nonexhaust + idw\_pm25monitoring\_stations +  linesnear + industry1000 + roads300 + ldres1000 + roads1000 + avg\_traffic\_stations + zbe + lat + lon \\

@@

PM\textsubscript{2.5 - Zn} & 2021 & 141 & 0.44 & 27.7 & pm25\_road\_nonexhaust + pm25\_background + idw\_pm25monitoring\_stations + linesnear + industry1000 + roads50 + build25 + roads300 + ldres1000 + roads1000 + avg\_traffic\_stations +  trafmajor +trafmajorload1000 +trafload1000 + hdres25 + zbe + lat + lon \\

@@

\small

@@

>>>>

\begin{enumerate}
    \item Define the baseline period, that is, the total period in which the air pollution measurements were done. For NO\textsubscript{2} the baseline period was from 23\textsuperscript{rd} October 2018 to 16\textsuperscript{th} February 2022.
    \item Collect hourly air pollution data from a routine monitoring site covering the baseline period. We used data from Palau Reial monitoring station, as this is the reference site for the BiSCAPE campaigns and the BiSC cohort. NO\textsubscript{2}, data at Palau Reial was missing for only 5 \% of the period. Even though the missing rate is low, we imputed missing data by doing a linear regression with NO\textsubscript{2} levels at the monitoring site of Sants (r = 0.84). After recovering those data, the percentage of missing data at Palau Reial decreased to 0.68 \%. 
    \item Calculate the average concentration of NO\textsubscript{2} for the routine monitoring site covering the baseline measurement period: C\textsubscript{PR}(avg).
    \item Calculate the concentrations of the pollutant at the routine monitoring site for the exact same period (t) as the sample collected at the BiSC and BiSCAPE locations in the different periods: CP\textsubscript{PR}(t). 
    \item Calculate for the routine monitoring site for each sample the difference between the concentration (C\textsubscript{PR}) and the average covering the baseline measurement period: D\_C\textsubscript{PR}(t) = C\textsubscript{PR}(t) - C\textsubscript{PR}(avg).
    \item Calculate for each sample at site i (C\textsubscript{i}(t)) the adjusted concentration (C\textsubscript{i, adj}(t)) by computing: C\textsubscript{i,adj}(t) = C\textsubscript{i}/D\_C\textsubscript{PR}(t).
\end{enumerate}

\subsubsection{PM\textsubscript{2.5}, BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn} deseasonalization}

For the PM\textsubscript{2.5}, BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn} the data available was obtained during the BiSCAPE campaigns. The \textit{ratio method} was used for deseasoanlisation. The method procedure is very similar to the difference method explained. The rationale for using a different method for these pollutants is that we aim to create LUR models for some PM\textsubscript{2.5} components. Since we do not have hourly data from the PM\textsubscript{2.5} components for the baseline period (data is only available as a daily average every four days), the deseasonalisation of these components needs to be done with total PM\textsubscript{2.5} conctrations. As the \textit{ratio method} is based on subtraction, this method cannot be applied to different pollutants than the reference. Samples collected within the BiSCAPE campaigns at Palau Reial showed lower concentrations than those collected online at Palau Reial. Using the ratio method in PM\textsubscript{2.5} concentrations may lead to over-subtracting.

\begin{enumerate}
  \item Define the baseline period, that is, the total period in which the air pollution measurements were done. For PM\textsubscript{2.5} and its chemical components and BC the baseline period is from 01/01/2021 to 16/02/2022. 
  \item Collect hourly air pollution data from a routine monitoring site covering the baseline period. We used data from Palau Reial monitoring site, as this is the reference site for the BiSC cohort. 
  \begin{enumerate}
      \item PM\textsubscript{2.5} data at Palau Reial was missing for 26.8\% of the period due instrument maintenance. We imputed missing data by doing a linear regression with collocated BC concentrations (r=0.57) as no other station in Barcelona had hourly data for PM\textsubscript{2.5} during this period. After recovering those data, the percentage of missing data at Palau Reial decreased to 1.0\%. 
      \item BC data at Palau Reial was missing for only 2.3\% of the period
  \end{enumerate}
 \item Calculate the average concentration of the corresponding pollutant (j) for the routine monitoring site covering the baseline measurement period: C\textsubscript{j,PR}(avg). For PM\textsubscript{2.5} components, in this step j corresponds to total PM\textsubscript{2.5} mass concentrations.
 \item Calculate the concentrations of the pollutant j at the routine monitoring site for the exact same period (t) as the sample collected at BiSCAPE locations in the different periods: C\textsubscript{j,PR}(t). For PM\textsubscript{2.5} components, in this step j corresponds to total PM\textsubscript{2.5} mass concentrations. 
 \item Calculate the routine monitoring site for each sample the ratio between the concentration of pollutant j (C\textsubscript{j, PR}(t)) and the average covering the baseline measurement period: R\_C\textsubscript{j,PR} = C\textsubscript{PR}(t) / C\textsubscript{PR}(avg). For PM\textsubscript{2.5} components, in this step j correspond to total PM\textsubscript{2.5} mass concentrations.
 \item Calculate for each sample at site i (C\textsubscript{i}(t)) the adjusted concentration of pollutant j (C\textsubscript{i,adj}(t)) by computing: C\textsubscript{j,i,adj}(t) = c\textsubscript{j,i}(t)/R\_C\textsubscript{j,PR}(t)
\end{enumerate}

\subsubsection{Averages for LUR model development}
Once the measurements done at different time points (t) and sites (i) for the different pollutants (j) were deseasonalised, we proceeded to calculate the average of the adjusted cocnentrations for each site i. Therefore, the average value obtained for each site was considered to be the average for the baseline period at site i. These averages were the input values for the development of the LUR model.

@@

Following the ESCAPE guidelines

@@

\subsubsection{Spatial variables}

%%% Figure: Monitoring stations and time-series used for LUR Back-Extrapolation %%%
\captionsetup[figure]{skip=6pt} % Adjusting the space between the figure and its caption
\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.6\textwidth]{figures/lur_pred_corr.png} % Adjust the path to your image
    \caption{Spatial location and daily temporal concentrations at Palau Reial station.}
    \label{FigureS1} % Label for referencing
\end{figure}

\subsubsection{Air pollution monitoring stations (XVPCA stations)}
%%% Figure: Monitoring stations and time-series used for LUR Back-Extrapolation %%%
\captionsetup[figure]{skip=6pt} % Adjusting the space between the figure and its caption
\begin{figure}[!htb]
    \centering
    \includegraphics[width=1.0\textwidth]{figures/location_ts_stations_hm.png} % Adjust the path to your image
    \caption{Spatial location and daily temporal concentrations at Palau Reial station.}
    \label{FigureS1} % Label for referencing
\end{figure}

\subsubsection{Traffic count measurement and meteorological data}
While our dispersion model initiall

\subsubsection{Dispersion model estimates}

@@

\clearpage
\begin{samepage}
\subsection{Model performance}
The performance metrics of the final HM for NO\textsubscript{2}, PM\textsubscript{2.5}, BC and PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn} is described in \textbf{\Cref{Table S4}}

@@

\begin{table}[!htbp]

@@

\label{Table S4}

@@

\small % Making the table text smaller

@@

 & & & & & avg\_traffic\_stations + avg\_solar\_radiation + \\ 

@@

\multirow{5}{*}{PM\textsubscript{2.5}} & \multirow{5}{*}{2021} & \multirow{5}{*}{161} & \multirow{5}{*}{0.66} & \multirow{5}{*}{3.45} & pm25\_dispersion + idw\_pm25monitoring\_stations + \\
& & & & & avg\_traffic\_stations + ldres1000 + build25 + \\
& & & & & pop1000 + roads25 + sqralt  + \\
& & & & &  majorroadlength25 + trafload300 + \\
& & & & & build\_height\_25 + lat + lon \\

@@

\multirow{6}{*}{BC} & \multirow{6}{*}{2021} &\multirow{6}{*}{74} & \multirow{6}{*}{0.86} & \multirow{6}{*}{0.23} & bc\_dispersion + idw\_nox\_monitoring\_stations +\\ 
& & & & & avg\_atmospheric\_pressure + avg\_bc\_palau\_reial +   \\ 
& & & & &  pop100 + hdres300 + roads25 + avg\_wind\_speed +  \\
& & & & & roadlength25 + hdres50 + distinvmajor1 +  \\
& & & & & avg\_traffic\_stations + linesnear + \\
& & & & & pop300 + trafload500 + roads500 \\

@@

\multirow{4}{*}{PM\textsubscript{2.5 - Fe}} & \multirow{4}{*}{2021} & \multirow{4}{*}{161}  & \multirow{4}{*}{0.54} & \multirow{4}{*}{0.08} & pm25\_road\_nonexhaust + \\ 
& & & & & idw\_pm25\_monitoring\_stations + \\
& & & & & roads300 + roads1000 + ldres500 + \\
& & & & & sqralt + zbe + lat + lon \\

@@

\multirow{5}{*}{PM\textsubscript{2.5 - Cu}} & \multirow{5}{*}{2021} & \multirow{5}{*}{154} & \multirow{5}{*}{0.70} & \multirow{5}{*}{2.15} & pm25\_road\_nonexhaust + \\
& & & & & idw\_pm25\_monitoring\_stations + \\
& & & & & linesnear + industry1000 + roads300 + \\
& & & & & ldres1000 + roads1000 + avg\_traffic\_stations + \\
& & & & &  zbe + lat + lon \\

@@

\multirow{6}{*}{PM\textsubscript{2.5 - Zn}} &\multirow{6}{*}{2021} & \multirow{6}{*}{141} & \multirow{6}{*}{0.44} & \multirow{6}{*}{27.7} & pm25\_road\_nonexhaust + pm25\_background + \\
& & & & & idw\_pm25monitoring\_stations + linesnear + \\
& & & & & industry1000 + roads50 + build25 + roads300 + \\
& & & & & ldres1000 + roads1000 + avg\_traffic\_stations + \\
& & & & & trafmajor + trafmajorload1000 +  trafload1000 +\\
& & & & & hdres25 + zbe + lat + lon \\

@@

\scriptsize

@@

\end{samepage}
\clearpage
